---
title: "cnidarian_meta-analysis_statistics"
author: "Nathan R. Geraldi"
date: "November 12, 2019"
output: github_document
---


Script is to run the statisical analysis of the cnidarian data
it uses the metafor package
It also estimates change with RCP porojections
And estimates what changes in temp or CO2 are needed for impacts


## libraries
```{r Libraries}

library(tidyverse)
library("metafor")
library("Rmisc")
library("plotrix")
library("tiff")

```

## functions
```{r_func eval=FALSE, include=FALSE}


```

## universal variables
```{r directory}
## file that has data
dat_file="/Users/nathangeraldi/Dropbox/Cnidaria meta-data/data/"
#  also in  dat_file="/Users/geraldn/Dropbox/Documents/KAUST/global change meta/R/export/"

## file to save data
save_file="/Users/nathangeraldi/Dropbox/Cnidaria meta-data/data/"
##  shared file to save data
save_file_shared="/Users/nathangeraldi/Dropbox/Cnidaria meta-data/data/"


```

## get data
meta-analysis data and others
```{r data}

dat<-read.table(paste(dat_file,"meta-analysis_used_data.csv",sep="" ), sep=",",header=T)   ## na

mhw<- read.table(paste(dat_file,"MHW projections.csv",sep="" ), sep=",",header=T)  
 # names(mhw) 
# names(dat)
```


## tidy data
keep only calcyfying corals and remove deep sea corals
```{r tidy_dat}

 sdat<- dat %>% #  names(sdat)   names(dat)   unique(dat$Deep.sea.)
   filter(Taxa_catagory=="Calcifying coral")  %>%  ##  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
   filter(Deep.sea.=="No" | is.na(Deep.sea.))  %>%  ##!!!!!!!!!!!!!!!!
  filter(Latitude.decimal.degrees <= 35) %>%   #!!!!!!!!!!!!!!!! remove a few norther studies
  #filter(Trait.type!= "growth") %>%   # !!!!!!!!!! remove growth - very few aftre corrections

   mutate(temp_change=Experimental_temp.level.degrees-Control_temp.level.degrees) %>% 
   mutate(co2_change=Experimental_pCO2.level.µatm-Control_pCO2.level.µatm) %>% # hist(sdat$co2_change)
     
   filter(co2_change <= 800 | is.na(co2_change)) %>% #  # !!!!!!!!!!!!!!!!limit to realistic changes in CO2
  
  ## added to clarify/simplity symbiont density -added Oct 2020  unique(dat$Trait.type)
  filter(! (Trait.type %in% "symbiont density"  & Trait.units %in% "% decrease in luminance ")  ) %>%
  filter(! (Trait.type %in% "symbiont density"  & Trait.units %in% "% corals bleached")  ) %>%
  filter(! (Trait.type %in% "symbiont density"  & Trait.units %in% "red channel intensity")  ) %>%
  filter(! (Trait.type %in% "symbiont density"  & Trait.units %in% "bleaching index (6-1)")  ) %>%
  # unique(sdat$Trait.units)
    mutate(Stressor.studied=factor(Stressor.studied)) %>% 
  mutate(Trait.type=factor(Trait.type)) %>% 
    droplevels()

  
mes<- sdat %>% 
  select(Stressor.studied,Trait.type,temp_AE) %>% 
  filter(Trait.type=="respiration" )
 

   
mhw<- mhw %>%   # names(mhw)
  mutate(Global.Warming_plus_mhw= Global.Warming_recent + Maximum.Annual.Intensity.of.Marine.Heatwaves)

```


## setup fro meta-analysis
```{r meta_setup}

# make blank dataframe to input stat results
op <- data.frame(Estimate=integer(),
                 SE=integer(),
                 zval=integer(),
                 pval=integer(),
                 ci.lb=integer(),
                 ci.ub=integer(),
                 stressor=character(),
                 variable=character(),
                 response=character(),
                 sample_size=integer(),
                 num_of_studies=integer(),
                 stringsAsFactors=FALSE)

#  set up variables and maximium AE
maxyi<-20  # code will remove the absolute value greater than this, removes very few but these are HUGE AE outlyers that make now logical sense

s_stud<-levels(sdat$Stressor.studied) 
resp<-levels(sdat$Trait.type)

#   names(sdat)
```

## meta-analysis
runs loop to run all category combinations - warming and/or CO2 manipulations (s_stud) and response trait measured (resp)
puts results into table
```{r meta-analysis}
###### start loop   ######
w<-1

for (j in c(1,2)) {   # for stressor
  
 # tryCatch({
  
    if (j==1) {
       sdat$yi<-sdat$temp_AE # !!!!!!! check which one  !!!!!!!!!!!
       sdat$vi<-sdat$temp_AE_variation
       sdat$change<-sdat$temp_change
       s_stud2 <-s_stud[c(3,2)]
       minchange<-0.5 # set this to filer out any temp change < minchange - difference between experimental and control
       qqq<-"temperature"
    } else {
       #sdat$yi<-sdat$AE_9eV_CO2ppm  #  class(sdat$yi)  names(sdat)
       sdat$yi<-sdat$CO2_AE
        sdat$vi<-sdat$CO2_AE_variation
        sdat$change<-sdat$co2_change
        s_stud2<-s_stud[1:2]
        minchange<-50  # set this to filer out any CO2 change < minchange - difference between experimental and control
        qqq<-"CO2"}
  
  for ( h in c(1,2)) {
      target<-s_stud2[h]

  for (i in c(1:length(resp))) {  # for diffreent catagories  i<-1
    #    target<-"warming"  #    names(sdat)
      sdat2 <- sdat %>% 
        filter(Stressor.studied==target & complete.cases(Stressor.studied)) %>% 
        filter(Trait.type==resp[i])  %>% 
        filter(change > minchange)  %>%    # 
        filter(abs(yi)<= maxyi) %>%  #   hist(sdat$AE_9eV_temp)

        dplyr::select(yi,vi,PROVINCE,Reference.ID,Genus) %>% 
        drop_na() %>% 
        filter(is.finite(vi))  %>% 
        filter(complete.cases(yi))  %>% 
        filter(!is.infinite(yi)) 
      
    #run model with each pub per species   
            m1<-rma.mv(yi,vi,random=list(~1 | Reference.ID, ~1 | Genus),
           data=sdat2, method="ML") 

            funnel(m1, level=c(95), shade=c("white", "gray", "darkgray"), refline=0)
            forest(m1)
            
            
        op[w,1]<-m1$beta  # ESTIMATE  q<-m1$beta[[1]]  op[1,1]<-q
        op[w,2]<-m1$se   # $SE
        op[w,3]<-m1$zval  # 
        op[w,4]<-m1$pval  # 
        op[w,5]<-m1$ci.lb   # 
        op[w,6]<-m1$ci.ub  # $S
        op[w,4]<-m1$pval  # 
        op[w,5]<-m1$ci.lb   # 
        op[w,6]<-m1$ci.ub  
        op[w,7]<-target
        op[w,8]<-resp[i]
        op[w,9]<-qqq
        op[w,10]<-length(row.names(sdat2))
        op[w,11]<-length(unique(sdat2$Reference.ID))

      w<-w+1
  }
  }
 #  }, error=function(e){})
}


#   output table is op
```



## meta diagnositcs
some code to check for biases and outliers

```{r meta_diag, eval=FALSE, include=FALSE}

sdat$yi<-sdat$AE_9eV_temp_pos_response # choose effect size metric
sdat$vi<-sdat$AE_9eV_temp_variation
sdat$change<-sdat$temp_change
s_stud2<-s_stud[c(3,2)]
minchange<-0.5 
qqq<-"temperature"
i<-3  # choose resp
h<-1  # choose warming - 1 or both - 2
target<-s_stud2[h]

sdat2 <- sdat %>%  # names(sdat2)
  filter(Stressor.studied==target & complete.cases(Stressor.studied)) %>% 
  filter(variable.label.fix.2.fix_PAM==resp[i])  %>% 
  filter(change > minchange)  %>%    # as suggested by Carlos
  filter(abs(yi)<= maxyi) %>%  #15?  25 works    hist(sdat$AE_9eV_temp)
  
  dplyr::select(yi,vi,PROVINCE,Original.ref.number,genus) %>% 
  drop_na() %>% 
  filter(is.finite(vi))  %>% 
  filter(complete.cases(yi))  %>% 
  filter(!is.infinite(yi)) %>% 
  group_by(Original.ref.number) %>% 
  mutate(study_num = as.character(match(Original.ref.number, unique(Original.ref.number))))  %>%
  ungroup() 


  
sdat2<- sdat2 %>% 
  filter(!(study_num==38))
 # filter(!(study_num==11 |study_num==46))
    

#run model with each pub per species    PROVINCE/
m1<-rma.mv(yi,vi,random=list(~1 | Original.ref.number, ~1 | genus),
           data=sdat2, method="ML")
m1


hist(m1$yi.f)
get_bad<-sdat2 %>% 
  mutate(mod_yi = m1$yi.f) %>% 
  filter (mod_yi>4)


funnel(m1, level=c(95), shade=c("white", "gray", "darkgray"), refline=0)
pp<-forest(m1)



## get summary
mes<-sdat2 %>%  # names(sdat2)
  group_by(Reference.ID) %>% 
  mutate(study_num = as.character(match(Reference.ID, unique(Reference.ID))))  %>%
  ungroup() %>% 
  group_by(study_num) %>% 
  dplyr::summarise(mean_st = mean(yi)) %>% 
  arrange(desc(mean_st))




```

# fix resp direction-old
```{r resp}
  # to make consistant for temp and respiration make ci.lb into ci.ub ,  unique(sdat$Trait.type)
#op<-op %>%   # names(op)
#    mutate(ci.lb= ifelse(response == "temperature"  & variable=="respiration" ,ci.ub , ci.lb) ) %>% 
 #   mutate(Estimate= ifelse(response == "temperature"  & variable=="respiration" ,abs(Estimate) , Estimate) )
```



## RCP projections

```{r rcp_proj}
## make table with rcp's and temp increase
rcp <- data.frame("RCP" = c(2.6,4.0,6.0,8.5), "RCP_temp" = c(1,1.8,2.2,3.7), "RCP_co2"=c(63,173,276,490)) # when using rcp means 
rcp_co<- data.frame("RCP" = c(2.6,4.5,6.0,8.5), "rcp_co2"=c(63,173,276,490)) 
# now using MHW data  names(mhw)   names(rcp)
rcp<- mhw %>% 
    filter( Percentile.of.SST == "90th") %>% 
    mutate(rcps=sub(" .*", "", rcps)) %>% 
    filter(rcps != "Historical") %>% 
    select(rcps, Year,Maximum.Annual.Intensity.of.Marine.Heatwaves,Global.Warming_plus_mhw) %>% 
    unite(rcp_years, rcps:Year, sep="_", remove=FALSE) %>% 
    pivot_longer(c(Maximum.Annual.Intensity.of.Marine.Heatwaves,Global.Warming_plus_mhw), names_to = "Projection_type", values_to = "projected_temp") %>% 
    unite(rcp_years_proj, rcps:Projection_type, sep="_", remove=FALSE) %>% 
    mutate(RCP=as.numeric(gsub("RCP","",rcps) ) ) %>%  # clean rcps remove "RCP"
    left_join(rcp_co) %>% # add in RCP - co2 add latter
    arrange(Projection_type, RCP, Year)


## calulate 
# names(rcp_proj)

#rcp_proj <- rcp[rep(op, times=length(rownames(rcp)) ) ,]

rcp_proj <-data.frame(op,i=rep(1:length(rownames(rcp)),ea=NROW(op)))
    
rcp_proj <-  data.frame(op,i=rep(1:length(rownames(rcp)),ea=NROW(op)))  %>%  # should be number of rcps?  op - out put of model
  dplyr::select (-i) %>% 
  mutate(rcp_years_proj=rep(rcp$rcp_years_proj, each=length(rownames(op)))) %>%  # add to then join, just do one step?
  left_join(rcp)  %>%   # join to rcp data
  mutate(rcp_change=if_else(response=="temperature",projected_temp,rcp_co2)) %>% # make change depending on variable-temp or co2
  mutate(CI=ci.ub-ci.lb) %>% # calc. confidence interval
  # to go from temp to AE used linear relationship between difference in 1/kt and difference in degree C use 0.128710787*(temp deg.C)-0.003138539
  mutate(relative_prop=if_else(response=="temperature", -0.128710787*projected_temp-0.003138539, rcp_co2/100)) %>%  # make relative to change
  mutate(mean_prop_estimate=exp(relative_prop*Estimate)) %>% # get projected mean estimate    exp(-0.56*0.06)  .928

  
  #calculate Confidence interval
 #  mutate(CI_estimate=((exp(ci.lb*relative_prop))-(exp(ci.ub*relative_prop)))/2) %>% # get project 95% CI previous old calc-bad
    mutate(CI_estimate=if_else(response=="temperature",( exp(ci.lb * -0.128710787  -0.003138539) ) -  (exp(Estimate * -0.128710787  -0.003138539  ) ), (exp(ci.ub) - (exp(Estimate)))  )) %>% 
   # respiration and growth have opposite response to change - no difference
  # mutate(CI_estimate=if_else(variable == "growth" | variable == "respiration" & response == "temperature", ( exp(ci.ub * -0.128710787  -0.003138539) ) -  (exp(Estimate * -0.128710787  -0.003138539  ) ), CI_estimate  )) %>% 
  # calc if CI overlaps with 0
  mutate(Significant=if_else( mean_prop_estimate+ abs(CI_estimate) < 1,"yes","no"))  %>%   # does 95% CI overlap with 0
  # respiration and growth have opposite response to change
  mutate(check=mean_prop_estimate - abs(CI_estimate)) %>% 
  mutate(Significant=if_else(variable == "respiration" & response == "temperature"  & check > 1,"yes", Significant))  %>%   # does 95% CI overlap with 0
    mutate(Significant=if_else(variable == "growth" & response == "temperature"  & check > 1,"yes", Significant))  %>% 
  select(-check) %>% 
    # remove extra co2 measures because only 3 rcps hsould = 388 rows
  unite(co2_remove, c(Estimate:sample_size,rcp_change,CI), sep="_", remove=FALSE) %>% 
  dplyr::distinct(co2_remove, .keep_all = TRUE)
  
### fix for resp and growth for temp




```


## setup estimate temp&CO2
create empty dataframe
```{r est_temp}
op2<-op %>%  # was op changed march   names(op2)
  mutate(temp_inc_mean_less_95=-1000, prop_mean_AE__ci_less_1=-1000, temp_inc_ci_less_1=-1000, temp_increment=-1000,prop_mean_AE=-1000, prop_CI_AE=-1000,  finding="None")
  # get mean AE that CI cross zero
  max_degree_change<-10  # should be 10
  s<-seq(0.01,max_degree_change,by=0.01)  # length(s)
  
 #temp_l<-length(s[s<=max_degree_change]) #length table should be if nothing removed
 #co2_l<-length(s[s>=1]) 
 #s_int<-seq(1,100,by=1)  # length(s_int)
dum <- data.frame("increment" = s+1000, "prop_mean_AE" = s+1000,"prop_CI_AE" = s+1000) 
``` 


## run estimate loop
calculate temperature or CO@ increase that CI no longer crosses 1 or is significant

```{r est_temp_run}

# c(1:length(op2$Estimate))
for (j in c(1:length(op2$Estimate)) ) {
  
  # create table of increase incremental change  - 1-1000 steps, increment i/100
  ## TO CHECK need to keep actual seq() in for loop, messes up when use variable name (s10)!!!
  for (i in seq(1,length(s),by=1)) {
   #  i<-43  j<-30
  ii<-i/100  # determine degree incremnt step - 
  ## add increase to AE estimate from model
  dum$increment[i]<-ii
  
  ## account for difference in sign between reponses
      if (op2$response[j] == "temperature"){
      ## get AE of increase (stepping by 0.1 degree C when ii=10 or ii - 0.001)
  KT_increm<- -0.128710787*ii-0.003138539
 # first_ci_old<-(  exp(op2$ci.lb[j] * -0.128710787*0.01  -0.003138539)  ) -  (    exp(op2$Estimate[j] * -0.128710787*  0.01  -0.003138539  )  )
  ## use original mean and CI from model to get CI in AE
    first_ci<-(  exp(op2$ci.lb[j] * -0.128710787  -0.003138539)  ) -  (    exp(op2$Estimate[j] * -0.128710787  -0.003138539  )  )
  
  dum$prop_mean_AE[i]<- exp(op2$Estimate[j] * KT_increm)
  # dum$prop_CI_AE[i]<- exp(op2$ci.lb[j] * KT_increm)  # lower bound for Temp - used before Mar 2020
  
  
  
 # exception for resp and growth 
      if (op2$variable[j] == "respiration" | op2$variable[j] == "growth"){
         dum$prop_CI_AE[i]<- dum$prop_mean_AE[i] - abs(first_ci) 
      } else {
  dum$prop_CI_AE[i]<- dum$prop_mean_AE[i] + abs(first_ci)  
      } # original diff + new mean_AE
  
    
  
  
  #dum$prop_CI_AE[i]<- (exp(  (op2$ci.ub[j]-op2$Estimate[j]) * KT_increm) ) + dum$prop_mean_AE[i] #  - bad
      } else {
        ## for CO2!!!
  dum$prop_mean_AE[i]<- exp(op2$Estimate[j] * ii) # increments relative to 100 atm (ie 110 uatm = 1.1)
   first_ci<-(exp(op2$ci.ub[j]) - (exp(op2$Estimate[j])))
  dum$prop_CI_AE[i]<- dum$prop_mean_AE[i] + first_ci #  upper bound for CO2
              }
  ### for checks, can remove
  #print(i): print(dum$increment[i]): print(KT_increm): print(dum$prop_mean_AE[i]): print(dum$prop_CI_AE[i])
  
  }  # end of dum creation loop
  
  
  ## created table now check for parameters
  ## first for temp then CO2
  
  ## for temp
  if (op2$response[j]=="temperature"){
    dum2<-dum[dum$increment<=max_degree_change,] ## limit findings to increments 0.1 to max
    temp_l<-length(dum2$increment)
     # find when prop_CI_ub_AE is <1, or approaches 1

    
     if (op2$variable[j] == "respiration" | op2$variable[j] == "growth"){
         less1<-dum2[dum2$prop_CI_AE >= 1  , ] #
      } else {
      less1<-dum2[dum2$prop_CI_AE <= 1  , ] # set cutoff- always 1 - was original code before adding exception above
      }
    
    
    
    #more1<-dum2[dum2$prop_CI_AE >= 1, ]
    
  # if prop_CI_AE becomes less than 1 get temp increment -- length(less1) <50 means nothing was removed from table
        if (length(less1$increment) == temp_l & length(less1$increment) > 0){
            outp<-less1[1,] # get first row when become <1
            answerp<-"temp alwasy significant, CI <=1"  # "temp_Significant change see CI for <=1"
        }
  
  
  # if prop_CI_AE is alwasy greater than 1 (length of orignial data) it is always significant, but then get temp. increase when prop_meanis < 0.95
        if (length(less1$increment) < temp_l) {
          
     #   less_mean_0.95n<- dum2[dum2$prop_mean_AE <= 0.95, ]
       if (op2$variable[j] == "respiration" | op2$variable[j] == "growth"){
         less_mean_0.95n<- dum2[dum2$prop_mean_AE >= 1.05, ] #
      } else {
      less_mean_0.95n<- dum2[dum2$prop_mean_AE <= 0.95, ] # set cutoff- always 1 - was original code before adding exception above
      }
     
            if (length(less_mean_0.95n$increment) < temp_l & length(less_mean_0.95n$increment) >=1) { # check to see if went below 95%
              outp<-less_mean_0.95n[1,]
              answerp<-"temp_Sig. CI , see incremnt mean <=95%"
            } else {                          # if never below 95%
              outp<-c(NA,NA,NA)
              answerp<-"temp_Sig. CI , 95% unknown"
                    }
          } # end of temp inc <
    
    # if prop_CI_AE always greater than 1 - not sensitive
        if (length(less1$increment) == 0) {

       
           if (op2$variable[j] == "respiration" | op2$variable[j] == "growth"){
         more_mean_0.95n<- dum2[dum2$prop_mean_AE >= 1.05, ] #
      } else {
      more_mean_0.95n<- dum2[dum2$prop_mean_AE <= 0.95, ] # set cutoff- always 1 - was original code before adding exception above
      }
       
       
            if (length(more_mean_0.95n$increment) < temp_l & length(more_mean_0.95n$increment) >=1) { # check to see if went below 95%
              outp<-more_mean_0.95n[1,]
              answerp<-"temp-CI always >1, increment mean <=95%"
            } else {                          # if never below 95%
              outp<-c(NA,NA,NA)
              answerp<-"temp-CI always >1, unknown 95%"
                    }
          } # end of temp inc equal
    
    #  get increment when mean prop become < or = 0.95
      
             if (op2$variable[j] == "respiration" | op2$variable[j] == "growth"){
         lessninefive<-dum[dum$prop_mean_AE >= 1.05  , ]  #
      } else {
      lessninefive<-dum[dum$prop_mean_AE <= 0.95  , ] # set cutoff- always 1 - was original code before adding exception above
      }
      
      
     if (length(lessninefive$increment) > 0){
            mean_95p<-lessninefive[1,1] # get first row when become <1
     } else {    
      mean_95p<-NA }
    
       #  get increment when ci prop becomes <= 1
      # set cutoff
         if (op2$variable[j] == "respiration" | op2$variable[j] == "growth"){
         less1_all<-dum[dum$prop_CI_AE >= 1  , ]  #
      } else {
       less1_all<-dum[dum$prop_CI_AE <= 1  , ] # set cutoff- always 1 - was original code before adding exception above
      }
      
      
     if (length(less1_all$increment) >= 1){
            ci_less_1<-less1_all[1,1] # get first row when become <1
              prop_when_ci_less_1<-less1_all[1,2]
     } else {    
            ci_less_1<-NA 
            prop_when_ci_less_1<-NA  }
    
    
  } #end of temp if
  ############################################################
  ##############################
  # for CO2 !!!!!!!!!!!!
  if (op2$response[j]=="CO2"){    
    dum2<-dum[dum$increment>=0,] ## limit findings to increments 1 to 10
    co2_l<-length(dum2$increment)
         # find when prop_CI_ub_AE is <1
  less1<-dum2[dum2$prop_CI_AE <= 1  , ]# set cutoff
  # if prop_CI_AE becomes less than 1 get temp increment -- length(less1)< co2_l means nothing was removed from table
   if (length(less1$increment) == co2_l & length(less1$increment) >0){
   outp<-less1[1,] # get first row when become <1
   answerp<-"CO2-Sig. change, see CI when it <=1"
      }
  
  # if prop_CI_AE is alwasy less than 1 (length of orignial data) it is always significant, but then get temp. increase when prop_meanis < 0.95
   if (length(less1$increment) < co2_l) {
     less_mean_0.95n<- dum2[dum2$prop_mean_AE <= 0.95, ]
     if (length(less_mean_0.95n$increment) < co2_l & length(less_mean_0.95n$increment) >0) { # check to see if went below 95%
   outp<-less_mean_0.95n[1,]
   answerp<-"CO2-Sig,CI always <1, mean became <=95%"
       } else {                          # if never below 95%
    outp<-c(NA,NA,NA)
    answerp<-"CO2-Sig.CI always <1, mean never"
      }
   } # end of inc
  
      # if prop_CI_AE alwasy greater than 1 - not sensitive
        if (length(less1$increment) == 0) {
       more_mean_0.95n<- dum2[dum2$prop_mean_AE <= 0.95, ]
            if (length(more_mean_0.95n$increment) < co2_l & length(more_mean_0.95n$increment) >0) { # check to see if went below 95%
              outp<-more_mean_0.95n[1,]
              answerp<-"co2-CI always >1, increment mean <=95%"
            } else {                          # if never below 95%
              outp<-c(NA,NA,NA)
              answerp<-"co2-CI always >1, unknown 95%"
                    }
          } # end of co2 inc equal
  
      #  get increment when mean prop become < or = 0.95
      lessninefive<-dum[dum$prop_mean_AE <= 0.95  , ] 
     if (length(lessninefive$increment) > 0){
            mean_95p<-lessninefive[1,1] # get first row when become <1
     } else {    
      mean_95p<-NA }
    
       #  get increment and AE prop when ci prop becomes <= 1
      less1_all<-dum[dum$prop_CI_AE <= 1  , ] # set cutoff
     if (length(less1_all$increment) > 0){
            ci_less_1<-less1_all[1,1] # get first row when become <1
            prop_when_ci_less_1<-less1_all[1,2]
     } else {    
            ci_less_1<-NA 
            prop_when_ci_less_1<-NA  }
  
  
  
  }  #end of co2 if
  
 ###### end of calc
  
  
    #out put to op2  names(op2)
  op2[j,(ncol(op2)-3):(ncol(op2)-1)]<-outp
  op2[j,ncol(op2)]<-answerp
  op2[j,ncol(op2)-4]<-ci_less_1
  op2[j,ncol(op2)-5]<-prop_when_ci_less_1
  op2[j,ncol(op2)-6]<-mean_95p
  
  ## clear outp
  outp<-c(222,222,222)
  ci_less_1<-222
  prop_when_ci_less_1<-222
  mean_95p<-222
} 

```


## export

```{r export}

#### export meta-analysis summary table to shared folder
    write.table(op, paste(save_file_shared,"meta-analysis_summary_oct2020_github.csv",sep=""), sep=",",row.names=F)

#### export rcp projection table to shared folder
    write.table(rcp_proj, paste(save_file_shared,"mhw_rcp_projections_from_meta_oct2020.csv",sep=""), sep=",",row.names=F)
    
    #### export meta-analysis summary table to shared folder
    op2.2 <- op2 %>% 
      select(-finding)
    
    write.table(op2.2, paste(save_file_shared,"meta-analysis_summary_oct2020_with_future.csv",sep=""), sep=",",row.names=F)
    
    
# export data used in analysis   - use   sdat
    write.table(sdat, paste(save_file_shared,"data_used_in_meta-analysis__oct2020.csv",sep=""), sep=",",row.names=F)
    
    
    
```



## map of studies

```{r map}
##########       map of corals with latitude frequnce      #############################
library(mapdata)
library(maps)

mdat<- sdat %>%   # names(sdat)
  mutate(unique_study_loc=paste(Reference.ID,Latitude.decimal.degrees, Longitude.decimal.degrees., sep="_")) %>% 
  filter(!duplicated(unique_study_loc))  %>%   # from 468 to 301
  dplyr::select(Latitude.decimal.degrees, Longitude.decimal.degrees.) 

par(fig=c(0,0.85,0,1), xpd=F)
maps::map("worldHires",xlim = c(-180,180), ylim = c(-85,85),fill=T,col="gray", border=F,
     cex.lab=1, mar=c(3,5,0,2))
map.axes(lty=1.2)
points(mdat$Longitude.decimal.degrees., mdat$Latitude.decimal.degrees, cex=.7)
#mtext("Longitude", SOUTH<-1, line=-5, adj=.42,outer=TRUE,cex=1)
mtext("Longitude", SOUTH<-1, line=-3.5, adj=.42,outer=TRUE,cex=1)
mtext("Latitude", WEST<-2, line=-1.2, adj=0.5,outer=TRUE,cex=1)
par(fig=c(0.83,1,0.237,0.765), mar=c(0,.5,0,1), new=TRUE)
h<-hist(mdat$Latitude.decimal.degrees,plot=F)
hh<-c(0,0,0,h$counts,0,0,0,0)  #  c(0,0,h$counts,0)
barplot(hh,width=1,horiz=T, beside=T,xlim=c(0,100),
        space=0, las=1)   # names.arg = hh$mids
axis(side=2, at=c(2,7,12), labels=c(-50,0,50),las=0)
box(lty=1.2)
#mtext("Locations", SOUTH<-1, line=-5.0, adj=.95,outer=TRUE,cex=1)
mtext("Locations", SOUTH<-1, line=-3.5, adj=.95,outer=TRUE,cex=1)

## save   ------  map_of_sites_may_2020


```






